# Reproducible Research: Peer Assessment1

## Loading and preprocessing the data
```{r echo=TRUE}
theData <- read.csv('/home/rodney/Projects/Git_repos/RepData_PeerAssessment1/activity.csv', stringsAsFactors=FALSE, header=TRUE)
#omit any NA's:
theData2 <- na.omit(theData)
#aggregate steps for each date:
totalSteps <- aggregate(steps~date,theData2,sum)
```

## Histogram of the total number of steps taken each day:
```{r}
hist(totalSteps$steps, col=1,main='Total steps taken each day',xlab='Total steps')
```

## What is mean and median total number of steps taken per day?
```{r}
#library(plyr)
#ddply(theData2, .(date),summarize, mean=mean(steps, na.rm=TRUE), median=median(steps, na.rm=T))

mean(totalSteps$steps)
median(totalSteps$steps)
```

## What is the average daily activity pattern?
```{r}
# aggregrate steps for each interval to get mean number of steps per interval
intervalSteps<- aggregate(steps ~ interval, theData2, mean)
#plot the line graph:
plot(intervalSteps$interval,intervalSteps$steps,type='l', col=1,main='Average number of steps for all days',xlab='Interval',ylab='Average number of steps')

#find row id of of max average number of steps
masRow <- which.max(intervalSteps$steps)
#find which interval has the max avg number of steps
intervalSteps[masRow,]
```
Maximum average number of steps is 206 found in interval 835.

## Imputing missing values
```{r}

```

## Are there differences in activity patterns between weekdays and weekends?
```{r}

```

